---
title: "Ologies_Analysis"
output:
  github_document:
    html_preview: false
date: "2024-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(RColorBrewer)
library(patchwork)
library(tidyr)
library(dplyr)
library(tidyverse)
library(stringr)


```


# What's in an Ology: A Breakdown of Ologies Episodes by Category

## Data wrangling

The dataframe read in here was generated using a python script which can be found in the github. This script was used to parse ologies episodes and extract meaningful information - duration of the episode, type of ology, and date.

Data was then manually curated to add in broad classifications and subcategories. These were based on Wikipedia designations of branches of science: Formal, Natural, and Social sciences; with Applied or Foundational designations. This was further broken down into additional branches. Since many episodes are interdisciplinary, we gave primary and secondary designations.

```{r}
df <- read.csv("Ologies_Data.csv")
View(df)

df <- df[-c(12:14)] #trimming out extraneous columns

ologies <- df[!is.na(df$ology),] #selecting just the episodes with ologies
ologies <- ologies[!is.na(ologies$branch),] #excluding any NAs for branch - this also excludes smologies
View(ologies)

#we want to exclude encores so they aren't counted twice

ologies <- subset(ologies, encore == 0)

#same with part 2 episodes
ologies <- subset(ologies, part_2 == 0)

#final dataframe

dim(ologies)
write.csv(ologies, file = "trimmed_ologies.csv")

ologies <- read.csv(file = "trimmed_ologies.csv")
```
 Final analysis includes 246 unique episodes of Ologies by Alie Ward.
 
## What is the most common primary topic of Ologies?
 

 
```{r}

#good but we can do better. let's reorder the dataframe so we can order by occurence


primary2 <- ggplot(data = ologies, aes(primary, fill = primary)) +
  geom_bar(stat = "count", col = "#d397fa", alpha = 0.2)  +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle("Ologies episode by primary topic")
  
  
primary2
```
 
 ## By broad category
 
```{r}
branch <- ggplot(data = ologies, aes(branch, fill = branch)) +
  geom_bar(stat = "count", col = "#d397fa", alpha = 0.2)  +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle("Ologies episode by branch of science")
  
  
branch
```
 
 Transform frequency into a count variable using dplyr
 
```{r}
ologies_count <- ologies %>%
  group_by(primary) %>%
  count(primary)
```

Re-order

```{r}
ologies_count <- as.data.frame(ologies_count)
colnames(ologies_count) <- c("Primary", "Count")

ologies_order <- ologies_count[order(ologies_count$Count),]
ologies_order$Primary <- factor(ologies_order$Primary, levels = ologies_order$Primary)

ggplot(ologies_order, aes(x = Primary, y = Count, fill = Primary)) + 
  geom_bar(stat = "identity", col = "#d397fa", alpha = 0.2)  +
  theme_minimal() +
  theme(axis.text.x = element_text(hjust = 1, vjust = 0.5)) +
  ggtitle("Ologies episode by primary scientific category")+
  coord_flip()
```

## Episode duration distribution

```{r}
head(ologies)

ologies_ms_order <- ologies[order(ologies$duration_ms),]
ologies_ms_order$duration_ms <- factor(ologies_ms_order$duration_ms, levels = ologies_ms_order$duration_ms)

ologies_ms_order <- subset(ologies_ms_order, X != 158)
ologies_ms_order <- subset(ologies_ms_order, X != 323)


durationplot <- ggplot(data = ologies_ms_order, aes(x = reorder(primary, -duration_min), y = duration_min, fill = primary)) +
  geom_boxplot(col = "#d397fa", alpha = 0.2) +
  geom_point(col = "#d397fa", alpha = 0.6)+
  theme_minimal() +
  xlab("Primary category") +
  ylab("Episode duration (minutes)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(legend.position = "none") +
  ggtitle("Duration of Ologies episodes")

durationplot
ggsave("duration_plot.jpg", plot = durationplot)
  
```

### Mean duration with error
```{r}

ggplot(ologies_ms_order, aes(x = reorder(primary, -duration_min), y = duration_min, fill = primary)) + 
  geom_bar(stat="summary", fun.y="mean", col = "#d397fa", alpha = 0.2) + 
geom_errorbar(stat="summary", colour = "#d397fa", width = 0.2,
				fun.ymin=function(x) {mean(x)-sd(x)/sqrt(length(x))}, 
				fun.ymax=function(x) {mean(x)+sd(x)/sqrt(length(x))}) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle("Mean duration of ologies episodes (with error)")

```


## Ologies over time: distribution by year, month

```{r}

#split the release date column so we can look by month, year

ologies[c('Month', 'Day', 'Year')] <- str_split_fixed(ologies$release_date, '/*/', 3)


year <- ggplot(data = ologies, aes(Year, fill = Year)) +
  geom_bar(stat = "count", col = "#d397fa", alpha = 0.2)  +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle("How many episodes released per year?")
  
year

```
### Month


### Months with correct labels

```{r}
#create month labels

monthNames2 <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

month <- ggplot(data = ologies, aes(Month, fill = Month)) +
  geom_bar(stat = "count", col = "#d397fa", alpha = 0.2)  +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(legend.position = "none")+
  scale_x_discrete(limits = c('1', '2', '3','4','5','6','7','8','9','10','11','12'), labels = monthNames2)+
  ggtitle("Frequency by month (cumulative)")
  
month

```

### Split by year

```{r}
#split by year

monthNames2 <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

monthyear <- ggplot(data = ologies, aes(Month, fill = Month)) +
  geom_bar(stat = "count", col = "#d397fa", alpha = 0.2)  +
  theme_minimal() +
  facet_wrap(~Year) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(legend.position = "none")+
  scale_x_discrete(limits = c('1', '2', '3','4','5','6','7','8','9','10','11','12'), labels = monthNames2)+
  ggtitle("Frequency by month and year")
  
monthyear

```



 